### 2014-12-05 writing/F'14 paper/discussion.tex
Paper update
Last paper update for now.

This paper first evaluates two recursive models on a series of three increasingly
challenging interpretive tasks involving natural language inference---the 
core relational algebra of natural logic with entailment and
exclusion; recursive propositional logic structures; and statements
involving quantification and negation. We then showed that the same models can learn to
perform an entailment task on natural language image captions. \mynote{The results suggest that RNTNs,
and potentially also RNNs}, have the capacity to model these tasks with 
reasonably-sized training sets. These positive results are
promising for the future of learned representation models in the
applied modeling of compositional semantics.

@@

Our artificial data experiments have only scratched the surface 
of the logical complexity of natural language; in future experiments, we hope to test sentences
with embedded quantifiers, multiple interacting quantifiers, relative
clauses, and other kinds of recursive structure. Nonetheless, the
rapid progress the field has made with these models in recent years
provides ample reason to be optimistic that they can be trained to
meet all the challenges of natural language semantics.

>>>>

This paper first evaluates two recursive models on a series of three increasingly
challenging interpretive tasks involving natural language inference---the 
core relational algebra of natural logic with entailment and
exclusion; recursive propositional logic structures; and statements
involving quantification and negation. We then show that the same models can learn to
perform an entailment task on natural language image captions. The results suggest that RNTNs,
and potentially also RNNs, have the capacity to learn to model these tasks with 
reasonably-sized training sets. These positive results are
promising for the future of learned representation models in the
applied modeling of compositional semantics.

@@

Our artificial data experiments have only scratched the surface 
of the logical complexity of natural language; in future experiments, we hope to test sentences
with embedded quantifiers, multiple interacting quantifiers, relative
clauses, and other kinds of recursive structure. Our SICK experiments 
similarly only begin to reveal the potential of these models to learn to 
perform complex semantic inferences from corpora. Nonetheless, the
rapid progress the field has made with these models in recent years
provides ample reason to be optimistic that they can be trained to
meet all the challenges of natural language semantics.

### 2014-12-05 writing/F'14 paper/intro.tex
Last paper update for now.

Supervised recursive neural network models (RNNs) for sentence meaning
have been successful in an array of sophisticated language tasks,
including sentiment analysis \cite{socher2011semi,irsoydeep},
image description \cite{sochergrounded}, and paraphrase detection
\cite{Socher-etal:2011:Paraphrase}. These results are encouraging for
the ability of these models to learn compositional semantic grammars.
However, it remains an open question whether they can achieve the
high-fidelity distributed representations of recent algebraic
approaches \cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,Hermann-etal:2013,rocktaschellow},
and whether they can match the performance of grammars based in logical forms
\cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}
when it comes to the core semantic concepts of quantification,
entailment, and contradiction needed to identify the relationships
between sentences like \ii{every animal walks}, \ii{every turtle
 moves}, and \ii{most reptiles don't move}.

@@

In our first set of experiments, we use this pre-specified logical
grammar to generate controlled data sets encoding the semantic
relationships between pairs of expressions and evaluate whether each
of two classes of neural model---plain recursive NNs and recursive neural
tensor networks (RNTNs; \citealt{socher2013acl1})---can learn those
relationships correctly. In our first experiment
(Section~\ref{sec:join}), we evaluate the ability of these models to
learn the core relational algebra of natural logic from data. Our
second experiment (Section~\ref{sec:recursion}) extends this
evaluation to relations between complex recursive structures like
$(a \plor b)$ and $\plneg (\plneg a \pland \plneg b)$, and our third
experiment (Section~\ref{sec:quantifiers}) involves relations between
statements with quantification and negation like \ii{every reptile walks} and \ii{no turtle
 moves}. \mynote{Though the performance of the plain RNN model 
 is somewhat poor in our first experiment, we find that the stronger RNTN
  model generalizes well in every case, suggesting that it has in fact
  learned, or at least learned to simulate, our target logical
  concepts.}

The experiments with simulated data provide a convincing demonstration
of the ability of neural networks to learn to encode semantic
inferences in complex natural language sentences from reasonably-sized
training sets. However, we are also interested in the more practical
question of whether they can learn these representations from
naturalistic text. To address this question, we apply our models to
the SICK entailment challenge data. The small size of this corpus puts
data-intensive neural models like ours at a disadvantage, but we are
nonetheless able to achieve competitive performance on it,
\mynote{surpassing several models with significant hand-engineered task-specific features.}
This suggests that the representational abilities that we observe in the previous sections
are not limited to carefully circumscribed tasks. We conclude that RNTN models
are adequate for typical cases of natural logic inference, and that there is not yet
any clear level of inferential complexity for which other approaches work and NN models fail.

>>>>

Supervised recursive neural network models (RNNs) for sentence meaning
have been successful in an array of sophisticated language tasks,
including sentiment analysis \cite{socher2011semi,irsoydeep},
image description \cite{sochergrounded}, and paraphrase detection
\cite{Socher-etal:2011:Paraphrase}. These results are encouraging for
the ability of these models to learn compositional semantic grammars.
However, it remains an open question whether they can achieve the
high-fidelity distributed representations of recent algebraic
approaches \cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,Hermann-etal:2013,rocktaschellow},
and whether they can match the performance of grammars based in logical forms
\cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}
on the core semantic issues of quantification,
entailment, and contradiction needed to identify the relationships
between sentences like \ii{every animal walks}, \ii{every turtle
 moves}, and \ii{most reptiles don't move}.

@@

In our first set of experiments, we use this pre-specified logical
grammar to generate controlled data sets encoding the semantic
relationships between pairs of expressions and evaluate whether each
of two classes of neural model---plain recursive NNs and recursive neural
tensor networks (RNTNs; \citealt{socher2013acl1})---can learn those
relationships correctly. In our first experiment
(Section~\ref{sec:join}), we evaluate the ability of these models to
learn the core relational algebra of natural logic from data. Our
second experiment (Section~\ref{sec:recursion}) extends this
evaluation to relations between complex recursive structures like
$(a \plor b)$ and $\plneg~(\plneg a \pland \plneg b)$, and our third
experiment (Section~\ref{sec:quantifiers}) covers relations between
statements with quantification and negation like \ii{every reptile walks} and \ii{no turtle
 moves}. Though the performance of the plain RNN model 
 is somewhat poor in our first experiment, we find that the stronger RNTN
  model generalizes well in every case, suggesting that it has in fact
  learned, or at least learned to simulate, our target logical
  concepts.

The experiments with simulated data provide a convincing demonstration
of the ability of neural networks to learn to encode semantic
inferences in complex natural language sentences from reasonably-sized
training sets. However, we are also interested in the more practical
question of whether they can learn these representations from
naturalistic text. To address this question, we apply our models to
the SICK entailment challenge data. The small size of this corpus puts
data-intensive neural models like ours at a disadvantage, but we are
nonetheless able to achieve competitive performance on it,
surpassing several models with significant hand-engineered task-specific features.
This suggests that the representational abilities that we observe in the previous sections
are not limited to carefully circumscribed tasks. We conclude that RNTN models
are adequate for typical cases of natural logic inference, and that there is not yet
any clear level of inferential complexity for which other approaches work and NN models fail.

### 2014-12-05 writing/F'14 paper/join.tex
Last paper update for now.

If a model is to learn the behavior of a relational logic like the one
presented here from a finite amount of data, it must learn to deduce new
relations from seen relations in a sound manner. The simplest such
deductions involve atomic statements using the relations in
Table~\ref{b-table}. For instance, given that $a \natrev b$ and $b
\natrev c$, one can conclude that $a \natrev c$, by basic
set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$.  The
full set of sound inferences of this form is summarized in
Table~\ref{tab:jointable}.

@@

\paragraph{Experiments}
To test our models' ability to learn these inferential patterns, we
create small boolean structures for our logic in which terms denote
sets of entities from a small domain.  Figure~\ref{lattice-figure}
depicts a structure of this form. The lattice gives the full model,
for which all the statements on below are valid. We divide these
statements evenly into training and test sets, and remove from the
test set statements which cannot be proven from the training
statements.
% using the logic depicted in Figure~\ref{lattice-figure}.

@@

We trained versions of both the RNN model and the RNTN model on these
data sets. In both cases, the models are implemented exactly as
described in Section~\ref{methods}, but since the items being compared
are single terms rather than full tree structures, the composition
layer is not involved, and the two models differ only in which
layer function is used for the comparison layer. We simply present
the models with two single terms, each of which corresponds to a
single vector in the (randomly initialized) vocabulary matrix $V$,
thereby ensuring that the model has no information about the terms
being compared except the relations between them.

\begin{figure}[t]
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \newcommand{\labelednode}[4]{\put(#1,#2){\oval(1.5,1)}\put(#1,#2){\makebox(0,0){$\begin{array}{c}#3\\\{#4\}\end{array}$}}}
    \setlength{\unitlength}{1cm}\scalebox{0.7}{
    \begin{picture}(5,5.5)
      \labelednode{2.50}{5}{}{1,2,3}

@@

\paragraph{Results} 
Table \ref{joinresultstable} shows the results, which are straightforward 
to interpret. The RNTN is able to accurately encode the relations 
between the terms in the geometric relations between their vectors, 
and is able to then use that information to recover relations that 
are not overtly included in the training data. The RNN, in contrast, was not able to learn
to do this well under these conditions, though we cannot rule
out the possibility that different optimization techniques or further hyperparameter 
tuning could lead an RNN model to succeed here.

@@

>>>>

If a model is to learn the behavior of a relational logic like the one
presented here from a finite amount of data, it must learn to deduce new
relations from seen relations in a sound manner. The simplest such
deductions involve atomic statements using the relations in
Table~\ref{b-table}. For instance, given that $a \natrev b$ and $b
\natrev c$, one can conclude that $a \natrev c$, by basic
set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$.  The
full set of sound such inferences is summarized in
Table~\ref{tab:jointable}.

@@

\paragraph{Experiments}
To test our models' ability to learn these inferential patterns, we
create small boolean structures for our logic in which terms denote
sets of entities from a small domain.  Figure~\ref{lattice-figure}
depicts a structure of this form. The lattice gives the full model,
for which all the statements below it are valid. We divide these
statements evenly into training and test sets, and remove from the
test set statements which cannot be proven from the training
statements.
% using the logic depicted in Figure~\ref{lattice-figure}.

@@

We trained versions of both the RNN model and the RNTN model on these
data sets. In both cases, the models are implemented as
described in Section~\ref{methods}, but since the items being compared
are single terms rather than full tree structures, the composition
layer is not involved, and the two models differ only in which
layer function is used for the comparison layer. We simply present
the models with the (randomly initialized) embedding vectors for each
of two terms, ensuring that the model has no information about the terms
being compared except for the relations between them that appear in training.

\begin{figure}[t]
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \newcommand{\labelednode}[4]{\put(#1,#2){\oval(1.5,1)}\put(#1,#2){\makebox(0,0){$\begin{array}{c}#3\\\{#4\}\end{array}$}}}
    \setlength{\unitlength}{1cm}\scalebox{0.8}{
    \begin{picture}(5,5.5)
      \labelednode{2.50}{5}{}{1,2,3}

@@

@@

\paragraph{Results} 
Table \ref{joinresultstable} shows the results. 
The RNTN is able to accurately encode the relations 
between the terms in the geometric relations between their vectors, 
and is able to then use that information to recover relations that 
are not overtly included in the training data. The RNN also generalizes fairly well, 
but makes enough errors that it remains an open question whether 
it is capable of learning representations with these properties. 
We cannot rule out the possibility that different optimization techniques or
further hyperparameter tuning could lead an RNN model to succeed here.

### 2014-12-05 writing/F'14 paper/methods.tex
Last paper update for now.

To apply recursive models to this inference class, we propose the tree 
pair model architecture depicted in Figure~\ref{sample-figure}, which is novel but
closely modeled after past work on recursive NNs. 
In it, the two phrases being compared are processed separately using a pair 
of recursive networks that share a single set of parameters. 
The resulting vectors are fed into a separate comparison
layer that is meant to generate a feature vector (fixed at 75 dimensions) capturing the
relation between the two phrases. The output of this layer is then
given to a softmax classifier, which in turn produces a hypothesized
distribution over the seven relations represented in Table~\ref{b-table}.

@@

% Removed 'minibatch' -> For two of the experiments, we use minibatches
% of size 1, which doesn't really count.
We train the model using stochastic gradient descent (SGD)
with learning rates computed using AdaDelta \cite{zeiler2012adadelta}.
Training times on CPUs vary from hours to days across experiments, and
the hyperparameters of AdaDelta, the dimensions of the representations,
and the initialization strategy are all tuned manually.
Because the classes are not balanced in the artificial data experiments, we report performance
using both accuracy and macroaveraged F1, since the latter emphasizes
 performance on infrequent classes. We compute macroaveraged F1 
as the harmonic mean of average precision and average recall, both computed
for all classes for which there is test data, setting precision to 0 
where it is not defined. On the artificial data experiments, we report mean
results over five fold cross-validation, though differences across runs are small.
Source code and generated data will be released after the review period.

>>>>

To apply these recursive models to our task, we propose the tree 
pair model architecture depicted in Figure~\ref{sample-figure}, which is novel but
closely modeled after past work on recursive NNs. 
In it, the two phrases being compared are processed separately using a pair 
of recursive networks that share a single set of parameters. 
The resulting vectors are fed into a separate comparison
layer that is meant to generate a feature vector (fixed at 75 dimensions) capturing the
relation between the two phrases. The output of this layer is then
given to a softmax classifier, which in turn produces a hypothesized
distribution over the seven relations represented in Table~\ref{b-table}.

@@

% Removed 'minibatch' -> For two of the experiments, we use minibatches
% of size 1, which doesn't really count.
We train the model using stochastic gradient descent (SGD)
with learning rates computed using AdaDelta \cite{zeiler2012adadelta}.
Training times on CPUs vary from hours to days across experiments, and
the hyperparameters of AdaDelta, the dimensions of the representations,
and the initialization strategy are all tuned manually.
Because the classes are not balanced in the artificial data experiments, we report performance
using both accuracy and macroaveraged F1, since the latter emphasizes
 performance on infrequent classes. We compute macroaveraged F1 
as the harmonic mean of average precision and average recall, both computed
for all classes for which there is test data, setting precision to 0 
where it is not defined. On the artificial data experiments, we report mean
results over five fold cross-validation. Variance across runs is typically no
more than two percentage points.
Source code and generated data will be released after the review period.

### 2014-12-05 writing/F'14 paper/quantifiers.tex
Last paper update for now.

We have seen that recursive NNs can learn an approximation of propositional
logic.  However, natural languages can express functional meanings of
considerably greater complexity than this.  As a step towards
investigating whether our models can capture this complexity, we now
directly measure the degree to which they are able to
develop suitable representations for the semantics of natural language
quantifiers like \ii{most} and \ii{all} as they interact with negation. Quantifiers 
and negation are far from
the only place in natural language where complex functional meanings
are found, but they are natural starting point, since they can be tested
in sentences whose structures are otherwise quite simple, and since they have
formed a standard case study in prior formal work on natural
language inference \cite{Icard:Moss:2013:LILT}.

@@

We evaluate the models by sampling 20\% of the data to test on and 
training on the remainder. In this setting, each model must learn a
complete reasoning system for the limited language and logic presented
in the training data. To succeed on this task, the model must be able
to recognize all of the lexical relations between the nouns, verbs,
and quantifiers and how they interact. For instance, it might see
\eqref{p1} and \eqref{p2} in training and be required to then label \eqref{p3}.

>>>>

We have seen that recursive NNs can learn an approximation of propositional
logic.  However, natural languages can express functional meanings of
considerably greater complexity than this.  As a step towards
investigating whether our models can capture this complexity, we now
directly measure the degree to which they are able to
develop suitable representations for the semantics of natural language
quantifiers like \ii{most} and \ii{all} as they interact with negation. Quantification 
and negation are far from
the only place in natural language where complex functional meanings
are found, but they are natural starting point, since they can be tested
in sentences whose structures are otherwise quite simple, and since they have
formed a standard case study in prior formal work on natural
language inference \cite{Icard:Moss:2013:LILT}.

@@

In each run, we sample 20\% of the 60k generated pairs to test on and 
train on the remainder. In this setting, each model must learn a
complete reasoning system for the limited language and logic presented
in the training data, and must be able
to recognize all of the lexical relations between the nouns, verbs,
and quantifiers and how they interact. For instance, it might see
pairs like \eqref{p1} and \eqref{p2} in training and be required to 
then label \eqref{p3}.

### 2014-12-05 writing/F'14 paper/recursion.tex
Last paper update for now.

\begin{table}[tp]
  \centering\small
  \begin{subtable}[t]{0.45\textwidth}
    \centering
    \begin{tabular}[t]{l l}
      \toprule
      Formula     & Interpretation \\
      \midrule
      $a$, $b$, $c$, $d$, $e$, $f$ & $\sem{x} \in \{\True, \False\}$ \\
      $\plneg \varphi$ & $\True$ iff $\sem{\varphi} = \False$ \\
      $(\varphi \pland \psi)$ & $\True$ iff $\False \notin \{\sem{\varphi}, \sem{\psi}\}$ \\
      $(\varphi \plor \psi)$  & $\True$ iff $\True \in \{\sem{\varphi}, \sem{\psi}\}$ \\
      \bottomrule
    \end{tabular}    
    \caption{Well-formed formulae. $\varphi$ and $\psi$
      range over all well-formed formulae, and $\sem{\cdot}$ is
      the interpretation function mapping formulae into $\{\True,
      \False\}$.}\label{tab:pl}
  \end{subtable}
  \begin{subtable}[t]{0.45\textwidth}
    \centering\vspace{4mm}
    \begin{tabular}[t]{r c l}
      \toprule
      $\plneg a$        & $\natneg$ & $a$ \\
      $\plneg \plneg a$ & $\nateq$  & $a$ \\
      $a$               & $\natfor$ & $(a \plor b)$ \\
      $a$               & $\natrev$ & $(a \pland b)$ \\
      %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
      $\plneg(\plneg a \pland \plneg b)$ & $\nateq$ & $(a \plor b)$ \\ 
      \bottomrule
    \end{tabular}
    \caption{Examples of statements about relations between
      well-formed formulae, defined in terms of sets of satisfying
      interpretation functions $\sem{\cdot}$.}\label{tab:plexs}
  \end{subtable}
  \caption{Natural logic relations over sentences of propositional logic.}  
  \label{prop-figure}
\end{table}

@@

\begin{figure*}[t]
  \centering
  \includegraphics[width=5in]{decayfig.eps}
%  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
  \caption{\mynote{Model performance on propositional logic, by expression size. }
%    The dashed line indicates that only expressions of size four or less appeared in the training data. \textbf{Bottom:} Semantically distinct formulae should have different
%    representations. As measured by Euclidean distance, only the RNTN
%    achieves this for formulae containing more than a small number of
%    connectives (\ii{and} in this example). The RNN quickly collapses the representations,
%    failing to capture the meaning contrasts.
    }  
  \label{prop-results}
  % TODO: Regenerate figure from final model.
\end{figure*}

\paragraph{Results} 
We trained both the RNN and RNTN models on data with up to 4
connectives (60k pairs) and tested it on a separate set of examples of up to 12
connectives (21k pairs). We initialized the model parameters randomly,
including the vector representations of the six variables.
Figure~\ref{prop-results} shows test the relationship
between test accuracy and statement length. \mynote{We found that both 
models were able to perform well on unseen small test examples, 
with RNN accuracy above
97\% and RNTN accuracy above 99\% on formulae below length five.
Starting at size four, performance gradually falls with increasing
size. Training accuracy was 99.4\% for the RNN and 99.9\% for the RNTN.}

The performance of these models on small unseen test examples
indicates that they learned correct approximations of the underlying
logic. It appears that these approximations are accurate enough to yield
correct answers when the composition layer is only applied a small
number of times, but that the error in the approximation grows with
increasing expression size, resulting in gradually dropping performance. 
It appears that the RNTN model overfit somewhat relative to the plain RNN,
performing better on the training data and short examples, but worse on longer
examples. Two factors are likely involved in this: even with the lower dimension,
the RNTN composition function has about eight times as many parameters as the
RNN, and this much more capacity to memorize a large number of narrow 
generalizations that work on the test data but don't generalize as well. In addition,
the RNTN used here used slightly less L2 regularization than the 
RNN ($\lambda_{RNN} = 0.001$, $\lambda_{RNTN} = 0.0003$), as had
yielded better performance earlier in tuning.

>>>>

\begin{table}[tp]
  \centering\small
  \begin{subtable}[t]{0.45\textwidth}
    \centering
    \begin{tabular}[t]{l l}
      \toprule
      Formula     & Interpretation \\
      \midrule
      $a$, $b$, $c$, $d$, $e$, $f$ & $\sem{x} \in \{\True, \False\}$ \\
      $\plneg \varphi$ & $\True$ iff $\sem{\varphi} = \False$ \\
      $(\varphi \pland \psi)$ & $\True$ iff $\False \notin \{\sem{\varphi}, \sem{\psi}\}$ \\
      $(\varphi \plor \psi)$  & $\True$ iff $\True \in \{\sem{\varphi}, \sem{\psi}\}$ \\
      \bottomrule
    \end{tabular}    
    \caption{Well-formed formulae. $\varphi$ and $\psi$
      range over all well-formed formulae, and $\sem{\cdot}$ is
      the interpretation function mapping formulae into $\{\True,
      \False\}$.}\label{tab:pl}
  \end{subtable}
  \begin{subtable}[t]{0.45\textwidth}
    \centering\vspace{4mm}
    \begin{tabular}[t]{r c l}
      \toprule
      $\plneg a$        & $\natneg$ & $a$ \\
      $\plneg \plneg a$ & $\nateq$  & $a$ \\
      $a$               & $\natfor$ & $(a \plor b)$ \\
      $a$               & $\natrev$ & $(a \pland b)$ \\
      %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
      $\plneg\, (\plneg a \pland \plneg b)$ & $\nateq$ & $(a \plor b)$ \\ 
      \bottomrule
    \end{tabular}
    \caption{Examples of statements about relations between
      well-formed formulae, defined in terms of sets of satisfying
      interpretation functions $\sem{\cdot}$.}\label{tab:plexs}
  \end{subtable}
  \caption{Natural logic relations over sentences of propositional logic.}  
  \label{prop-figure}
\end{table}

@@

\begin{figure*}[t]
  \centering
  \includegraphics[width=5.75in]{decayfig.eps}
%  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
  \caption{Model performance on propositional logic, by expression size.}
%    The dashed line indicates that only expressions of size four or less appeared in the training data. \textbf{Bottom:} Semantically distinct formulae should have different
%    representations. As measured by Euclidean distance, only the RNTN
%    achieves this for formulae containing more than a small number of
%    connectives (\ii{and} in this example). The RNN quickly collapses the representations,
%    failing to capture the meaning contrasts.

  \label{prop-results}

\end{figure*}

\paragraph{Results} 
We trained both the RNN and RNTN models on data with up to 4
connectives (60k pairs) and tested it on a separate set of examples of up to 12
connectives (21k pairs). We initialized the model parameters randomly,
including the vector representations of the six variables.
Figure~\ref{prop-results} shows test the relationship
between test accuracy and statement length. We found that both 
models were able to perform well on unseen small test examples, 
with RNN accuracy above
98\% and RNTN accuracy above 99\% on formulae below length five.
Starting at size four, performance gradually falls with increasing
size. Training accuracy was 99.4\% for the RNN and 99.8\% for the RNTN.

The performance of these models on small unseen test examples
indicates that they learned correct approximations of the underlying
logic. These approximations are accurate enough to yield
correct answers when the composition layer is only applied a small
number of times, but that the error in the approximation grows with
increasing expression size, resulting in gradually dropping performance. 
It appears that the RNTN model overfit somewhat relative to the plain RNN,
performing better on the training data and short examples, but worse on longer
examples. Two factors are likely involved in this: even with the lower dimension,
the RNTN composition function has about eight times as many parameters as the
RNN, and thus more capacity to find a solution involving a large number of narrow 
generalizations that work on the training data but don't generalize well. In addition,
the RNTN used here has slightly weaker L2 regularization than the 
RNN ($\lambda_{\textit{RNN}} = 0.001$, $\lambda_{\textit{RNTN}} = 0.0003$). This
yielded better average performance, but may have facilitated this effect.

### 2014-12-05 writing/F'14 paper/sick.tex
Last paper update for now.

The tree pair model architecture that we use is novel, and though the underlying recursive approach has been validated elsewhere, there is no guarantee that this architecture is suitable for handling inference in the face of the noisy labels and diverse range of linguistic structures seen in typical natural language data, and our experiments on artificial data are not alone sufficient to address this issue. To investigate the ability of the model to learn on natural data, we train a version of our model on the SICK textual entailment challenge corpus \cite{marelli2014sick}. The corpus consists of about 10k natural language sentence pairs (generated using data expansion from a smaller set), labeled with \ii{entailment}, \ii{contradiction}, or \ii{neutral}. At only a few thousand distinct sentences (many of them variants on an even smaller set of template sentences), the corpus is not large enough to train a high quality learned model of general natural language, but it is the largest hand labeled entailment corpus that we are aware of.

\begin{table*}[htp]
  \centering\small
  \begin{tabular}{lcl}
    \toprule
A woman is not boiling shrimps.& \ii{contradiction}&	A woman is boiling shrimps.\\
A wild deer is jumping a fence. &\ii{entailment}	&A deer is jumping a fence.\\
A potato is being sliced by a woman. &\ii{entailment}	&A woman is slicing a potato.\\
The plane, which is south African, is flying in a blue sky.& \ii{entailment}&	An airplane is flying through the air.\\
    \bottomrule
  \end{tabular}
  \caption{\label{examplesofsickdata}Examples of successful classifications on SICK.}
\end{table*}

Adapting to this task required us to make a few additions to the techniques discussed in Section \ref{methods}. In order to better handle rare words, we initialized our word embeddings using the 200 dimensional vectors trained with 
GloVe \cite{pennington2014glove} on data from Wikipedia. Since 200 dimensional vectors are too large to be practical in an RNTN on a small dataset, a new embedding transformation layer is needed. Before any embedding is used as an input to a recursive layer, it is passed through an additional $\tanh$ neural network layer with the same output dimension as the recursive layer which aggregates any usable information from the embedding vectors into a more compact working representation.

We also supplemented the SICK training data\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the SICK test set. We also report training accuracy on a small sample from each data source.} with 600k examples of entailment data from the Denotation Graph project (DG, \citealt{hodoshimage}), a corpus of automatically labeled (and thus noisy) entailment examples over image captions, the same genre of text from which SICK was drawn. We trained a single model on data from both sources, but used a separate set of softmax parameters for classifying into the labels from each source. We parsed the data from both sources with the Stanford PCFG Parser v. 3.3.1 \cite{klein2003accurate}. We also found that we were able to train a working model much more quickly with an additional technique: we collapse subtrees that were identical across both sentences in a pair down to a single head word. The test data on which we report performance are collapsed in this way, and both collapsed and uncollapsed copies of the training data are used in training. Finally, in order to improve regularization on the noisier data, we used dropout \cite{srivastava2014dropout} at the input to the comparison layer (10\%) and at the output from the embedding transform layer (25\%).

\begin{table}[tp]
  \centering \small
  \begin{tabular}{ l r@{ \ } r@{ \ } r@{ \ } }
    \toprule
    ~&\multicolumn{1}{c}{\ii{neut.} only} & \multicolumn{1}{c}{30d RNN}  & \multicolumn{1}{c}{50d RNTN}\\
    \midrule
    SICK Train &  56.7 & 95.6 &  \textbf{97.8}  \\
    DG Train &  50.0 & 68.0 & \textbf{74.0}  \\
    Test & 56.7 & 75.3 & \textbf{76.9}  \\
    \bottomrule
  \end{tabular}
  \caption{\mynote{Performance on SICK.}}
  \label{sresultstable}
\end{table}

The results are shown in Table \ref{sresultstable}. Despite the small amount of high quality training data available and the lack of resources for learning lexical relationships (especially exclusion, as in \ii{cat}--\ii{dog}), it is possible to train our model to perform competitively on textual entailment. Our performance did not reach that of the winning system (84.6\%), but exceeded the performance of eight out of 18 submitted systems including several which used sophisticated hand-engineered features and lexical resources specific to the version of the entailment task at hand.

Table \ref{examplesofsickdata} shows a few typical examples of correct classifications. The model is able to successfully recognize negation and the insertion and deletion of words, as demonstrated in the first two examples. The latter two show that the model can recognize entailments and contradictions between superficially different sentences, such as passive--active pairs and sentences with low lexical overlap.

>>>>

The tree pair model architecture that we use is novel, and though the underlying recursive approach has been validated elsewhere, there is no guarantee that this architecture is suitable for handling inference in the face of the noisy labels and diverse range of linguistic structures seen in typical natural language data, and our experiments on artificial data are not alone sufficient to address this issue. To investigate the ability of the model to learn on natural data, we train a version of our model on the SICK textual entailment challenge corpus \cite{marelli2014sick}. The corpus consists of about 10k natural language sentence pairs, labeled with \ii{entailment}, \ii{contradiction}, or \ii{neutral}. At only a few thousand distinct sentences (many of them variants on an even smaller set of template sentences), the corpus is not large enough to train a high quality learned model of general natural language, but it is the largest human-labeled entailment corpus that we are aware of.

\begin{table*}[htp]
  \centering\small
  \begin{tabular}{lcl}
    \toprule
A woman is not boiling shrimps.& \ii{contradiction}&	A woman is boiling shrimps.\\
A wild deer is jumping a fence. &\ii{entailment}	&A deer is jumping a fence.\\
A potato is being sliced by a woman. &\ii{entailment}	&A woman is slicing a potato.\\
The plane, which is south African, is flying in a blue sky.& \ii{entailment}&	An airplane is flying through the air.\\
    \bottomrule
  \end{tabular}
  \caption{\label{examplesofsickdata}Examples of correct RNTN classifications on SICK.}
\end{table*}

Adapting to this task required us to make a few additions to the techniques discussed in Section \ref{methods}. In order to better handle rare words, we initialized our word embeddings using 200 dimensional vectors trained with 
GloVe \cite{pennington2014glove} on data from Wikipedia. Since 200 dimensional vectors are too large to be practical in an RNTN on a small dataset, a new embedding transformation layer is needed. Before any embedding is used as an input to a recursive layer, it is passed through an additional $\tanh$ neural network layer with the same output dimension as the recursive layer. This new layer aggregates any usable information from the embedding vectors into a more compact working representation.

We also supplemented the SICK training data\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the SICK test set. We also report training accuracy on a small sample from each data source.} with 600k examples of entailment data from the Denotation Graph project (DG, \citealt{hodoshimage}), a corpus of automatically labeled (and thus noisy) entailment examples over image captions, the same genre of text from which SICK was drawn. We trained a single model on data from both sources, but used a separate set of softmax parameters for classifying into the labels from each source. We parsed the data from both sources with the Stanford PCFG Parser v. 3.3.1 \cite{klein2003accurate}. We also found that we were able to train a working model much more quickly with an additional technique: we collapse subtrees that were identical across both sentences in a pair down to a single head word. The training and test data on which we report performance are collapsed in this way, and both collapsed and uncollapsed copies of the training data are used in training. Finally, in order to improve regularization on the noisier data, we used dropout \cite{srivastava2014dropout} at the input to the comparison layer (10\%) and at the output from the embedding transform layer (25\%).

\begin{table}[tp]
  \centering \small
  \begin{tabular}{ l r@{ \ } r@{ \ } r@{ \ } }
    \toprule
    ~&\multicolumn{1}{c}{\ii{neut.} only} & \multicolumn{1}{c}{30d RNN}  & \multicolumn{1}{c}{50d RNTN}\\
    \midrule
    SICK Train &  56.7 & 95.4 &  \textbf{97.8}  \\
    DG (Train) &  50.0 & 67.0 & \textbf{74.0}  \\
    SICK Test & 56.7 & 74.9 & \textbf{76.9}  \\
    \bottomrule
  \end{tabular}
  \caption{Performance on collapsed SICK data.}
  \label{sresultstable}
\end{table}

The results are shown in Table \ref{sresultstable}. Despite the small amount of high quality training data available and the lack of resources for learning lexical relationships (especially exclusion, as in \ii{cat}--\ii{dog}), it is possible to train our model to perform competitively on textual entailment. Our RNTN did not reach the performance of the winning system (84.6\%), but it did exceed that of eight out of 18 submitted systems, including several which used sophisticated hand-engineered features and lexical resources specific to the version of the entailment task at hand.

Table \ref{examplesofsickdata} shows a few typical examples of correct classifications. Our qualitative evaluation suggests that the model is able to successfully recognize negation and the insertion or deletion of words, as in the first two examples. The latter two show that the model can also identify the relationships between superficially different sentences, such as passive--active pairs and sentences with low lexical overlap.
